A Generic and Efficient E-field Parallel Imaging Correlator for Next-Generation Radio Telescopes

### 2015-11-06 imaging_demo/ms.tex
Update section 1 with changes suggested by Adam
From a01e819 (master):
=====================

In ms.tex,

(*) Update section 1 with changes suggested by Adam

Radio astronomy is entering an era in which interferometers of hundreds to
thousands of individual antennas are needed to achieve desired survey speeds.
Nowhere is this more apparent than at radio frequencies below 1.4 GHz. The study
of the history of hydrogen gas throughout the universe's evolution is pushing
technology development towards arrays of low-cost antennas with large fields of
view and densely packed apertures. Similarly, the search for transient objects
and regular monitoring of the time-dependent sky is driving instruments in the
same direction with the added requirement of fast read-outs. A number of new 
telescopes are under development around the world based on this new paradigm, 
including the Murchison Widefield Array (MWA;\citealt{tin13,bow13}), the 
Precision Array for Probing the Epoch of Reionization (PAPER; \citealt{par10}), 
the Hydrogen Epoch of Reionization Array\footnote{http://reionization.org} 
(HERA), the LOw Frequency ARray (LOFAR;\citealt{van13}), the Canadian Hydrogen 
Intensity Mapping Experiment (CHIME,\citealt{ban14}), the Long Wavelength 
Array (LWA, \citealt{ell13}), and the low frequency Square Kilometer Array 
(SKA1-Low \citealt{mel13}).

@@

For certain classes of radio arrays there is an alternative to the FX correlator
that can lower the computational burden by directly performing a spatial fast
Fourier transform (FFT) on the electric fields measured by each antenna in the
array at each time step, removing the cross-correlation step. This relieves the
computational scaling from the harsh $\Nant^2$ to the more gentle envelope of
$\Ngrid\log_2\Ngrid$, where $\Ngrid$ is the number of pixels in the Fourier 
transform \citep[e.g.][]{mor11,teg09,teg10}. This architecture is often referred 
to as a ``direct imaging" correlator because it eliminates the intermediary 
cross-correlation data products of the FX and earlier lag correlators, but 
instead directly forms images from the electric field measurements.

@@

We begin with a technical description of the algorithm in \S\ref{sec:math}, 
then discuss our particular implementation in \S\ref{sec:software}. We then 
verify the output data quality from our code in \S\ref{sec:verify} by presenting 
simulated images from both the EPIC correlator and comparing to a simulated FX 
correlator. We also demonstrate the performance with real-world data from the 
LWA. In \S \ref{sec:analysis} we analyze the scaling relationships of the 
algorithm. We identify specific array design classes where the EPIC correlator 
is computationally more efficient and in the field of transients, demands 
significantly lesser bandwidth relative to visibility based approaches. We 
conclude and discuss future research prospects in \S\ref{sec:conclusions}.

>>>>

Radio astronomy is entering an era in which interferometers of hundreds to
thousands of individual antennas are needed to achieve desired survey speeds.
Nowhere is this more apparent than at radio frequencies below 1.4 GHz. The study
of the history of hydrogen gas throughout the universe's evolution is pushing
technology development towards arrays of low-cost antennas with large fields of
view and densely packed layouts. Similarly, the search for transient objects
and regular monitoring of the time-dependent sky is driving instruments in the
same direction with the added requirement of fast read-outs. A number of new 
telescopes are being or were developed around the world based on this new 
paradigm, the Hydrogen Epoch of Reionization 
Array\footnote{http://reionization.org} (HERA), including the Murchison 
Widefield Array (MWA;\citealt{tin13,bow13}), the Precision Array for Probing 
the Epoch of Reionization (PAPER; \citealt{par10}), the LOw Frequency ARray 
(LOFAR;\citealt{van13}), the Canadian Hydrogen Intensity Mapping Experiment 
(CHIME,\citealt{ban14}), the Long Wavelength Array (LWA, \citealt{ell13}), and 
the low frequency Square Kilometer Array (SKA1-Low \citealt{mel13}).

@@

For certain classes of radio arrays there is an alternative to the FX correlator
that can lower the computational burden by directly performing a spatial fast
Fourier transform (FFT) on the electric fields measured by each antenna in the
array at each time step, removing the cross-correlation step. This relieves the
computational scaling from the harsh $\Nant^2$ to the more gentle envelope of
$\sim\Ngrid\log_2\Ngrid$, where $\sim\Ngrid$ is the number of grid points in the 
Fourier transform \citep[e.g.][]{mor11,teg09,teg10}. This architecture is often 
referred to as a ``direct imaging'' correlator because it eliminates the 
intermediate cross-correlation data products of the FX and XF correlators, but 
instead directly forms images from the electric field measurements.

@@

We begin with a technical description of the algorithm in \S\ref{sec:math}, 
then discuss our particular implementation in \S\ref{sec:software}. We then 
verify the output data quality from our code in \S\ref{sec:verify} by presenting 
simulated images from both the EPIC correlator and comparing to a simulated FX 
correlator. We also demonstrate the performance with real-world data from the 
LWA. In \S \ref{sec:analysis}, we explore the scalability of the algorithm in the 
context of several array design choices. We identify specific array design 
classes where the EPIC correlator -- is computationally more efficient; and in 
the field of transients, demands significantly lesser I/O bandwidth relative to 
visibility based approaches. We conclude and discuss future research prospects 
in \S\ref{sec:conclusions}.
